\chapter{List of Papers}

\section*{Papers Included in this Thesis Proposal}
\begin{itemize}
    \item \textbf{Caitao Zhan}, Himanshu Gupta, Arani Bhattacharya, Mohammad Ghaderibaneh, "Efficient Localization of Multiple Intruders in Shared Spectrum System," IPSN, Sydney, Australia, Apr. 2020.

    \item \textbf{Caitao Zhan}, Mohammad Ghaderibaneh, Himanshu Gupta, "DeepMTL: Deep Learning Based Multiple Transmitter Localization," WoWMoM, Pisa, Italy, June. 2021. 

    \item \textbf{Caitao Zhan}, Mohammad Ghaderibaneh, Himanshu Gupta, "DeepMTL Pro: Deep Learning Based Multiple Transmitter Localization and Power Estimation," Pervasive and Mobile Computing\footnote{An extension of the WoWMoM conference paper}, 2022.

    \item \textbf{Caitao Zhan}, Himanshu Gupta, "Localize Transmitter in Quantum Sensor Networks", IEEE ICC 2022 (Under review)

    \item Mohammad Ghaderibaneh, \textbf{Caitao Zhan}\footnote{Only the balanced-tree section is included}, Himanshu Gupta, C.R. Ramakrishnan, "Efficient Quantum Network Communication using Optimized Entanglement Swapping Trees," IEEE TQE, 2022.


\end{itemize}

\section*{Other Papers}
\begin{itemize}

    \item Mark Hillery, Himanshu Gupta, \textbf{Caitao Zhan}, "Discrete Outcome Quantum Sensor Networks", Physical Review A (Under review)

    \item Mohammad Ghaderibaneh, \textbf{Caitao Zhan}, Himanshu Gupta, "DeepAlloc: CNN-Based Approach to Efficient Spectrum Allocation in Shared Spectrum Systems," (Under submission)

    \item Arani Bhattacharya, \textbf{Caitao Zhan}, Abhishek Maji, Himanshu Gupta, Samir R. Das, Petar M. Djuric, Selection of Sensors for Efficient Transmitter Localization, IEEE/ACM Transactions on Networking 2021.

    \item Himanshu Gupta, Max Curran, \textbf{Caitao Zhan}. Near-Optimal Multihop Scheduling in General Circuit-Switched Networks, ACM CoNEXT 2020
    
    \item Arani Bhattacharya, \textbf{Caitao Zhan}, Himanshu Gupta, Samir R. Das, Petar M. Djuric. Selection of Sensors for Efficient Transmitter Localization, IEEE INFOCOM 2020
  
\end{itemize}

\section*{Poster}
\begin{itemize}

    \item \textbf{Caitao Zhan}, Himanshu Gupta, Mark Hillery, "Discrete Outcome Quantum Sensor Networks"
    \footnote{The poster has the same name as a PRA paper under review, but the content is different, and will later be written as a new PRA paper}, Stony Brook University, Graduate Research Day 2022
  
\end{itemize}