We address the problem of localizing multiple intruders (unauthorized
transmitters) using a distributed set of sensors in the context of a
shared spectrum system. 
In contrast to single transmitter localization, multiple
transmitter localization (\mtl) has not been thoroughly studied.
In shared spectrum systems, it is important to be able to
localize simultaneously present multiple intruders to effectively
protect a shared spectrum from malware-based, jamming, or other
multi-device unauthorized-usage attacks. The key challenge in solving
the MTL problem comes from the need to ``separate'' an aggregated
signal received from multiple intruders into separate
signals from individual intruders. Furthermore, in a shared spectrum
paradigm, presence of an evolving set of authorized users (e.g.,
primary and secondary users) adds to the challenge.

In this paper, we propose an efficient algorithm for the \mtl problem
based on the hypothesis-based Bayesian approach called
\map. Direct application of the \map approach to
the \mtl problem incurs prohibitive computational and training
cost. In this work, we develop optimized techniques based on \map with
significantly improved computational and training costs. In
particular, we develop a novel interpolation method, \ildw, which helps
minimize the training cost. We generalize our techniques via
online-learning to the setting wherein there may be a set of
dynamically-changing authorized users present in the background.
We evaluate our developed techniques on large-scale simulations as
well as on small-scale indoor and outdoor testbeds. Our experiments
demonstrate that our technique outperforms the prior approaches by
significant margins,  i.e., error up to 74\% less in large-scale simulations and 30\% less in real-world testbeds.


