In this chapter, we address the problem of Multiple Transmitter Localization (\mtl).
\mtl is to determine the
locations of potential multiple transmitters in a field, based on readings from a distributed set of sensors.
In contrast to the widely studied single transmitter localization problem, the \mtl problem has only been studied recently in a few works.
\mtl is of great significance in many applications wherein intruders
may be present. E.g., in shared spectrum systems, detection of unauthorized transmitters and estimating their power are imperative to 
efficient utilization of the shared spectrum.
	
In this chapter, we present \our, a novel deep learning approach to address the \mtl problem.  
In particular, we frame
\mtl as a sequence of two steps, each of which is a computer vision problem: image-to-image translation and object detection. 
The first step of image-to-image translation essentially maps an input image representing sensor readings to an 
image representing the distribution of transmitter locations, and the second object detection step derives precise
locations of transmitters from the image of transmitter distributions.
%%%%%%%%%%%%%%%%%%%
For the first step, we design our learning model \imgimg, while for the second step, we customize a 
state-of-the-art object detection model \yolocust. 
%%%%%%%%%
Using  \our as a building block, we also develop techniques to estimate transmit power of the localized transmitters.
%%%%%%%%
We demonstrate the effectiveness of our approach
via extensive large-scale simulations and show that our approach outperforms the previous approaches 
significantly (by 50\% or more) in performance metrics including localization error, miss rate, and false alarm rate. 
Our method also incurs a very small latency.
We evaluate our techniques over a small-scale area with real testbed data and the testbed results align with the simulation results.
